---
title: "Plot The logistic model of Debowski"
author: "Uyen 'Rachel' Lai and Paul Sheridan"
output: html_notebook
---
```{r}

suppressWarnings(suppressMessages({
  library(tidyverse)
  library(jsonlite)
  library(scales)
  library(here)
}))


# Delete all variables initialized in R session, if any
rm(list = ls())

# Load libraries
library(tidyverse)
library(here)
library(reticulate)
library(magicaxis)
require(broom)
library(RColorBrewer)
library(scales)
library(latex2exp)
library(grid)

# Set plot colors 
gg_color_hue = function(n, alpha = 1) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100, alpha = alpha)[1 : n]
}
cols <- c("grey25", gg_color_hue(n = 3))
tcols <- c("grey25", gg_color_hue(n = 3, alpha = 0.5))
gray25 <- cols[1]; red <- cols[2]; green <- cols[3]; blue <- cols[4]
tred <- cols[2]; tgreen <- tcols[3]; tblue <- tcols[4]

```

#open vocabulary case

# The logistic model of Debowski params + config Json
```{r}
FILES_OPEN <- tribble(
  ~path,                                     ~label,
  "../../PubMed_human_Open.json",                 "Human",
  "../../PubMed_gptneo-125M_fewshot_Open.json",   "GPT-Neo 125M",
  "../../PubMed_gptneo-1.3B_fewshot_Open.json",   "GPT-Neo 1.3B",
  "../../PubMed_gptneo-2.7B_fewshot_Open.json",   "GPT-Neo 2.7B"
)

# Open-vocab G-model parameters
G_PARAMS_OPEN <- tribble(
  ~label,          ~alpha,    ~beta,    ~gamma,
  "Human",         -5.1689,   0.5427,   0.8353,
  "GPT-Neo 125M",  -2.7535,   0.6061,   1.0610,
  "GPT-Neo 1.3B",  -4.1221,   0.5521,   0.9545,
  "GPT-Neo 2.7B",  -4.6606 ,   0.5421,   0.8845
)

TOKEN_CAP_PER_SUBLIST <- 225
RANDOM_SEED           <- 42L

OUT_PDF <- here("pubmed-open-vocab-Gmodel.pdf")
```

# Assign color
```{r}
# Colours 
hex <- hue_pal()(3)
gg_red <- hex[1]
gg_green <- hex[2]
gg_blue <- hex[3]
gg_orange <- brewer.pal(n = 11, name = "PuOr")[4]
gg_purple <- "#C77CFF"

COLORS <- c(
"Human"     = gg_red,
"GPT-Neo 125M" = gg_blue,
"GPT-Neo 1.3B" = gg_green,
"GPT-Neo 2.7B" = gg_orange
)
corpora_levels <- c("Human", "GPT-Neo 125M", "GPT-Neo 1.3B", "GPT-Neo 2.7B")
names(COLORS) <- corpora_levels
```


# The logistic model of Debowski 
```{r}


g_paper <- function(n, alpha, beta, gamma) {
  n <- pmax(as.numeric(n), 1)
  nprime <- exp(-alpha) * n
  c <- 2 ^ ((1 - beta) / gamma)
  c * nprime / ((nprime^gamma + 1) ^ ((1 - beta) / gamma))
}
```

```{r}
cumulative_NV_from_sublists_fast <- function(docs,
                                             cap_per_sublist,
                                             seed = 42L) {
  set.seed(seed)

  idx  <- sample(seq_along(docs))
  docs <- docs[idx]

  capped <- lapply(docs, function(sub) {
    if (!is.list(sub) && !is.vector(sub)) return(character(0))
    sub <- unlist(sub, use.names = FALSE)
    if (length(sub) == 0) return(character(0))
    sub[seq_len(min(length(sub), cap_per_sublist))]
  })
  capped <- capped[lengths(capped) > 0]

  N_vec <- numeric(length(capped))
  V_vec <- numeric(length(capped))

  seen_env <- new.env(hash = TRUE, parent = emptyenv())
  N <- 0L
  V <- 0L

  for (i in seq_along(capped)) {
    sub <- capped[[i]]
    N   <- N + length(sub)

    for (tok in sub) {
      if (!exists(tok, envir = seen_env, inherits = FALSE)) {
        assign(tok, TRUE, envir = seen_env)
        V <- V + 1L
      }
    }

    N_vec[i] <- N
    V_vec[i] <- V
  }

  tibble(n = N_vec, V = V_vec)
}
```


# Build The logistic model of Debowski curves 

```{r}
raw_list   <- list()
g_list     <- list()

for (i in seq_len(nrow(FILES_OPEN))) {
  path  <- FILES_OPEN$path[i]
  label <- FILES_OPEN$label[i]

  if (!file.exists(path) && file.exists(paste0(path, ".json"))) {
    path <- paste0(path, ".json")
  }
  if (!file.exists(path)) {
    warning("File not found: ", path)
    next
  }

  message("Reading: ", path)
  docs <- jsonlite::fromJSON(path)

  # empirical open-vocab V(N), sublist-level
  nv <- cumulative_NV_from_sublists_fast(
    docs,
    cap_per_sublist = TOKEN_CAP_PER_SUBLIST,
    seed            = RANDOM_SEED
  )
  nv$corpus <- label
  raw_list[[label]] <- nv

  # G-model curve for this corpus
  pars <- G_PARAMS_OPEN %>% filter(label == !!label)
  if (nrow(pars) != 1) stop("No unique parameters for label ", label)

  alpha <- pars$alpha
  beta  <- pars$beta
  gamma <- pars$gamma

  n_min  <- max(1, min(nv$n))
  n_max_c <- max(nv$n)
  n_grid <- seq(n_min, n_max_c, length.out = 1000)
  V_grid <- g_paper(n_grid, alpha, beta, gamma)

  g_list[[label]] <- tibble(
    n      = n_grid,
    V      = V_grid,
    corpus = label
  )

  rm(docs)
  gc()
}

raw_open    <- bind_rows(raw_list)
g_data_open <- bind_rows(g_list)

raw_open$corpus    <- factor(raw_open$corpus,    levels = corpora_levels)
g_data_open$corpus <- factor(g_data_open$corpus, levels = corpora_levels)
```


# Legend label parameters 
```{r}


alpha_h   <- -5.1689; beta_h   <- 0.5418; gamma_h   <- 0.8353
alpha_125 <- -2.7535; beta_125 <- 0.6069; gamma_125 <- 1.0610
alpha_13  <- -4.1221; beta_13  <- 0.5461; gamma_13  <- 0.9545
alpha_27  <- -4.6606; beta_27  <-0.5421; gamma_27  <-  0.8845

n_max <- max(raw_open$n, na.rm = TRUE) *1.1
V_max <- max(raw_open$V, na.rm = TRUE) *1.1

scientific <- function(x) format(x, scientific = TRUE)
```

# Plot
```{r}

p <- ggplot()  +
  # G-model curves with color by corpus
  geom_line(
    data = g_data_open,
    aes(x = n, y = V, color = corpus),
    linewidth = 1.3
  ) +
  # empirical curves (all corpora) in gray
  geom_line(
    data = raw_open,
    aes(
      x = n,
      y = V,
      group = corpus,
      color = "Empirical Data"  
    ),
    linewidth = 0.6,
    linetype = 2
  )  +
  theme_minimal() +
  theme(
    legend.position      = c(0.99, 0.01),
    legend.justification = c("right", "bottom"),
    legend.text          = element_text(size = rel(0.75)),
    legend.background    = element_rect(fill = "gray95", color = NA),
    plot.margin          = unit(c(1, 1, 1, 1), "cm")
  ) +
  labs(
    title = "Open vocabulary setting",
    x     = "Corpus size",
    y     = "Vocabulary size",
    color = NULL
  ) +
  scale_color_manual(
    breaks = c("Empirical Data", names(COLORS)),
    values = c(
      "Empirical Data" = "grey35",
      COLORS
    ),
    labels = list(
      "Empirical data",
      bquote("PubMed Abstracts:" ~
             hat(alpha) == .(alpha_h)   ~ ", " ~
             hat(beta)  == .(beta_h)    ~ ", " ~
             hat(gamma) == .(gamma_h)),
      bquote("GPT-Neo 125M:" ~
             hat(alpha) == .(alpha_125) ~ ", " ~
             hat(beta)  == .(beta_125)  ~ ", " ~
             hat(gamma) == .(gamma_125)),
      bquote("GPT-Neo 1.3B:" ~
             hat(alpha) == .(alpha_13)  ~ ", " ~
             hat(beta)  == .(beta_13)   ~ ", " ~
             hat(gamma) == .(gamma_13)),
      bquote("GPT-Neo 2.7B:" ~
             hat(alpha) == .(alpha_27)  ~ ", " ~
             hat(beta)  == .(beta_27)   ~ ", " ~
             hat(gamma) == .(gamma_27))
    ),
    name = NULL
  )  +
  scale_x_continuous(
    limits = c(0, n_max),
    expand = expansion(mult = c(0, 0.05)),
    labels = scientific
  ) +
  scale_y_continuous(
    limits = c(0, V_max),
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        # 1 entry for "Empirical Data", then 4 corpora:
        linetype = c("dashed", "solid", "solid", "solid", "solid"),
        linewidth = c(0.6, 1.3, 1.3, 1.3, 1.3)
      )
    )
  )
p
```

# Save to PDF
```{r}

ggsave(
  plot      = p,
  filename  = OUT_PDF,
  device    = "pdf",
  width     = 6,
  height    = 6,
  units     = "in",
  dpi       = 320,
  limitsize = FALSE
)
```




# Closed vocabulary JSONs 

# The logistic model of Debowski params and Json config

```{r}



FILES_CLOSE <- tribble(
  ~path,                                     ~label,
  "../../data/PubMed_human_Close.json",                 "Human",
  "../../data/PubMed_gptneo-125M_fewshot_Close.json",   "GPT-Neo 125M",
  "../../data/PubMed_gptneo-1.3B_fewshot_Close.json",   "GPT-Neo 1.3B",
  "../../data/PubMed_gptneo-2.7B_fewshot_Close.json",   "GPT-Neo 2.7B"
)



# Closed-vocab G-model parameters (your numbers)
G_PARAMS_CLOSE <- tribble(
  ~label,          ~alpha,    ~beta,    ~gamma,
  "Human",         -10.1964,  0.1205,  0.0570,
  "GPT-Neo 125M",  -8.4498,   0.1432,  0.0617,
  "GPT-Neo 1.3B",  -9.0939,   0.1309,  0.0594,
  "GPT-Neo 2.7B",  -9.3873,   0.1285,  0.0590
)

TOKEN_CAP_PER_SUBLIST <- 225
RANDOM_SEED           <- 42L

OUT_PDF <- here("pubmed-closed-vocab-Gmodel.pdf")

```


# Colours
```{r}
# Custom color scheme
hex <- hue_pal()(3)
gg_red <- hex[1]
gg_green <- hex[2]
gg_blue <- hex[3]
gg_orange <- brewer.pal(n = 11, name = "PuOr")[4]
gg_purple <- "#C77CFF"




COLORS <- c(
"Human"     = gg_red,
"GPT-Neo 125M" = gg_blue,
"GPT-Neo 1.3B" = gg_green,
"GPT-Neo 2.7B" = gg_orange
)
corpora_levels <- c("Human", "GPT-Neo 125M", "GPT-Neo 1.3B", "GPT-Neo 2.7B")
names(COLORS) <- corpora_levels
```

# The logistic model of Debowski 
```{r}

g_paper <- function(n, alpha, beta, gamma) {
  n <- pmax(as.numeric(n), 1)
  nprime <- exp(-alpha) * n
  c <- 2 ^ ((1 - beta) / gamma)
  c * nprime / ((nprime^gamma + 1) ^ ((1 - beta) / gamma))
}
```
```{r}

cumulative_NV_from_sublists_fast <- function(docs,
                                             cap_per_sublist,
                                             seed = 42L) {
  set.seed(seed)

  idx  <- sample(seq_along(docs))
  docs <- docs[idx]

  capped <- lapply(docs, function(sub) {
    if (!is.list(sub) && !is.vector(sub)) return(character(0))
    sub <- unlist(sub, use.names = FALSE)
    if (length(sub) == 0) return(character(0))
    sub[seq_len(min(length(sub), cap_per_sublist))]
  })
  capped <- capped[lengths(capped) > 0]

  N_vec <- numeric(length(capped))
  V_vec <- numeric(length(capped))

  seen_env <- new.env(hash = TRUE, parent = emptyenv())
  N <- 0L
  V <- 0L

  for (i in seq_along(capped)) {
    sub <- capped[[i]]
    N   <- N + length(sub)

    for (tok in sub) {
      if (!exists(tok, envir = seen_env, inherits = FALSE)) {
        assign(tok, TRUE, envir = seen_env)
        V <- V + 1L
      }
    }

    N_vec[i] <- N
    V_vec[i] <- V
  }

  tibble(n = N_vec, V = V_vec)
}

```

# Build The logistic model of Debowski  curves
```{r}

raw_list   <- list()
g_list     <- list()

for (i in seq_len(nrow(FILES_CLOSE))) {
  path  <- FILES_CLOSE$path[i]
  label <- FILES_CLOSE$label[i]

  if (!file.exists(path) && file.exists(paste0(path, ".json"))) {
    path <- paste0(path, ".json")
  }
  if (!file.exists(path)) {
    warning("File not found: ", path)
    next
  }

  message("Reading: ", path)
  docs <- jsonlite::fromJSON(path)

  # empirical closed-vocab V(N), sublist-level
  nv <- cumulative_NV_from_sublists_fast(
    docs,
    cap_per_sublist = TOKEN_CAP_PER_SUBLIST,
    seed            = RANDOM_SEED
  )
  nv$corpus <- label
  raw_list[[label]] <- nv

  # G-model curve for this corpus
  pars <- G_PARAMS_CLOSE %>% filter(label == !!label)
  if (nrow(pars) != 1) stop("No unique parameters for label ", label)

  alpha <- pars$alpha
  beta  <- pars$beta
  gamma <- pars$gamma

  n_min <- max(1, min(nv$n))
  n_max_c <- max(nv$n)
  n_grid <- seq(n_min, n_max_c, length.out = 1000)
  V_grid <- g_paper(n_grid, alpha, beta, gamma)

  g_list[[label]] <- tibble(
    n      = n_grid,
    V      = V_grid,
    corpus = label
  )

  rm(docs)
  gc()
}

raw_close   <- bind_rows(raw_list)
g_data_close <- bind_rows(g_list)

raw_close$corpus   <- factor(raw_close$corpus,   levels = corpora_levels)
g_data_close$corpus <- factor(g_data_close$corpus, levels = corpora_levels)
```
  
  
# Legend label parameters 
  
```{r}

# Closed-vocab G-model parameters (your numbers)
G_PARAMS_CLOSE <- tribble(
  ~label,          ~alpha,    ~beta,    ~gamma,
  "Human",         -10.1964,  0.1205,  0.0570,
  "GPT-Neo 125M",  -8.4498,   0.1432,  0.0617,
  "GPT-Neo 1.3B",  -9.0939,   0.1309,  0.0594,
  "GPT-Neo 2.7B",  -9.3873,   0.1285,  0.0590
)

alpha_h   <- -10.1964; beta_h   <- 0.1205; gamma_h   <- 0.0570
alpha_125 <- -8.4498;  beta_125 <- 0.1432; gamma_125 <- 0.0617
alpha_13  <- -9.0939;  beta_13  <- 0.1309; gamma_13  <- 0.0594
alpha_27  <- -9.3873;  beta_27  <- 0.1285; gamma_27  <- 0.0590
n_max <- max(raw_close$n, na.rm = TRUE) * 1.1
V_max <- max(raw_close$V, na.rm = TRUE)* 1.1

scientific <- function(x) format(x, scientific = TRUE)
```

# Plot 
```{r}

p <- ggplot()  +
  geom_line(
    data = g_data_close,
    aes(x = n, y = V, color = corpus),
    linewidth = 1.3
  )+
  # empirical curves (all corpora) in gray
  geom_line(
    data = raw_close,
    aes(
      x = n,
      y = V,
      group = corpus,
      color = "Empirical Data"   # <--- put into aes() so it appears in legend
    ),
    linewidth = 0.6,
    linetype = 2
  )  +
  theme_minimal() +
  theme(
    legend.position      = c(0.99, 0.01),
    legend.justification = c("right", "bottom"),
    legend.text          = element_text(size = rel(0.75)),
    legend.background    = element_rect(fill = "gray95", color = NA),
    plot.margin          = unit(c(1, 1, 1, 1), "cm")
  ) +
  labs(
    title = "Closed vocabulary setting",
    x     = "Corpus size",
    y     = "Vocabulary size",
    color = NULL
  ) +
  scale_color_manual(
    # first "Empirical data", then your 4 corpora
    breaks = c("Empirical Data", names(COLORS)),
    values = c(
      "Empirical Data" = "grey35",  # gray for raw curves
      COLORS                        # your original colors for fits
    ),
    labels = c(
      "Empirical data",
      bquote("PubMed Abstracts:" ~
             hat(alpha) == .(alpha_h)   ~ ", " ~
             hat(beta)  == .(beta_h)    ~ ", " ~
             hat(gamma) == .(gamma_h)),
      bquote("GPT-Neo 125M:" ~
             hat(alpha) == .(alpha_125) ~ ", " ~
             hat(beta)  == .(beta_125)  ~ ", " ~
             hat(gamma) == .(gamma_125)),
      bquote("GPT-Neo 1.3B:" ~
             hat(alpha) == .(alpha_13)  ~ ", " ~
             hat(beta)  == .(beta_13)   ~ ", " ~
             hat(gamma) == .(gamma_13)),
      bquote("GPT-Neo 2.7B:" ~
             hat(alpha) == .(alpha_27)  ~ ", " ~
             hat(beta)  == .(beta_27)   ~ ", " ~
             hat(gamma) == .(gamma_27))
   ),
    name = NULL
  ) +
  scale_x_continuous(
    limits = c(0, n_max),
    expand = expansion(mult = c(0, 0.05)),
    labels = scientific
  ) +
  scale_y_continuous(
    limits = c(0, V_max),
    expand = c(0, 0)
  )+
  scale_x_continuous(
    limits = c(0, n_max),
    expand = expansion(mult = c(0, 0.05)),
    labels = scientific
  ) +
  scale_y_continuous(
    limits = c(0, V_max),
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        # 1 entry for "Empirical Data", then 4 corpora:
        linetype = c("dashed", "solid", "solid", "solid", "solid"),
        linewidth = c(0.6, 1.3, 1.3, 1.3, 1.3)
      )
    )
  )

p


```

# Save 
```{r}
ggsave(
  plot      = p,
  filename  = OUT_PDF,
  device    = "pdf",
  width     = 6,
  height    = 6,
  units     = "in",
  dpi       = 320,
  limitsize = FALSE
)


```